Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-12-02 leadership_across_cultures_contexts.Rmd

Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity. The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture. Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015]. Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample. We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework. We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity.

In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*. 2) Identify *features* of the qualities and functions of leaders (e.g., do particular variables tended to covary or cluster together?) in reference to theoretical models or novel interpretations. Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001]. 3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures. Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.

@@

<!-- Our fourth goal was to explore associations between the most common leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size. We used linear mixed effects models from the lme4 package [@bates2015], with common qualities and functions as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.  -->

Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups. We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.

>>>>

Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity. The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture. Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015]. Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample. We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework. We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity

In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*. 2) Identify *features* of the qualities and functions of leaders (e.g., do particular variables tended to covary or cluster together?) in reference to theoretical models or novel interpretations. Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001]. 3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers. Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.

@@

<!-- Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups. We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts. -->
